\documentclass{article}

%% This template is pieces of templates provided by Chi-Kit Lam for
%% CS395T in S2016, bits taken from
%% https://github.com/nykh/latex-shorthands/blob/master/myshorthands.tex, and
%% other configs scrapped together over time.

%% Settings for general styling, page layout, and such
\input{includes/layout.tex}

%% Settings for figures, captions, floats, tables, etc.
\input{includes/figs.tex}
%% Math settings
% Includes math packages, simple aliases for common symbols, and a few environments
\input{includes/mathstub.tex}
%% Section Separators
% Simple section separators, accessible with \starhsep and \clubhsep
\input{includes/separators.tex}
%% Color settings
% Include xcolors w/ svgnames, define a few special names
% Must be included before hyperlinks and listings
\input{includes/colors.tex}

%% Link settings
% Set sane default appearances for hyperlinked text
\input{includes/hyperlinks.tex}

%% Code snippet settings
% Sane default appearances for listing environments
\input{includes/listings.tex}

%% Other packages we need
\usepackage[version=4]{mhchem} % chemical formulae
\usepackage[numbers,square]{natbib} %Who uses non-numerical?

%%% style.sty ends%%%

\title{Graphics Final Report---Pathtracer}
\author{Kevin Song and Aditya Durvasula}
\date{\today}

\begin{document}
\maketitle

\section{Basic of Path Tracing}

Recall that the ultimate goal of almost all raytracing techniques is to simulate
the rendering equation, which is given by 

\begin{equation}
 L_{out}(\theta_r,\phi_r) = \int_{\theta_i} \int_{\phi_i} 
 f_r(\theta_r,\phi_r,\theta_i,\phi_i) L_{in}(\theta_i,\phi_i) \cos(\theta_i)
\end{equation}

where $f_r$ is the BRDF, encoding the properties of the material.

If we could find the solution for this equation directly, most of the rendering
problem would be solved. The problem with finding a solution is that the values 
for $L_{in}$ rely on the amount of light coming off of the surrounding materials.
To find the amount of light coming off of the surrounding materials, we need
to solve the rendering equation for those surfaces. Unfortunately, this means
solving even more copies of the rendering equation, and\dots.

The Whitted illumination model used in the ``standard'' raytracer solves this
by assuming point lights and direct illumination, i.e. only light that arrives
directly from the light contributes to $L_{in}$. While this is fast, it removes
things such as soft shadows, reflections, and specular highlights. While some
of these can be added back in, a truly globally correct illumnation model can
only be implemented with better calculations of the $L_{in}$ term.

\textit{Path tracing} attempts to solve the $L_{in}$ term in a global manner:
it attempts to integrate over the actual $L_{in}$ function by taking samples
and averaging their contribution. In essence, we make the following 
approximation:

\begin{equation}
    \int_{v \in \text{hemisphere}} L_{in}(v) \approx 
    \frac{1}{n} \sum_{i=0}^n L_{in}(v_i)
\end{equation}

where the $v_i$ are sampled randomly in the hemisphere.

We do a similar sampling trick to get soft shadows: instead of firing
a single shadow ray from the intersection towards a light, we randomly sample
points on the light, then fire many shadow rays from the intersection to the 
light. The actual value of the shadow at that point is the average of the
shadow samples.

\section{Implementation}

Our implementation uses a na\"ive MC sampler to integrate the rendering 
equation: a pair of random numbers on $[0,1)$ is turned into a random vector
on the unit sphere. This vector is then used to sample the hemisphere
and the contributions are averaged.

Interestingly, when we first implemented this, GLM's \texttt{sphericalRand()}
was proving to be a bottleneck. We implemented the sampler ourselves, only to
find that over 90\% of the program's runtime was being spent in calls to the
C++ random device. Given that this implementation uses an LCG, which is
generally considered the gold standard of fast, medicore-quality randomness,
this was very surprising. We eventually implemented our own fast, multithreaded
XorShift32 generator, which knocked random vector generation down to 60\% of
the program's runtime.

Our soft shadow implementation is similar to what is found in the extra credit
for this project--a certain number of samples are taken on the light, and shadow
rays are fired from the intersection to the light.

\end{document}